Assessment 1 Reproducible Research
========================================================

This is the Markdown document for the first Assignment of the course *Reproducible Research*

## Loading and preprocessing the data

This part is divided in two parts

1 For loading the data we use the following code

```{r}
datos <- read.csv("activity.csv",header=T)

```
2 The date was processed converting tha date column as Date class

```{r}
datos$date <- as.Date(datos$date)

```

## What is mean total number of steps taken per day?

For this part we ignore de NA's values  
1 For the histogram of total number of step taken each day we first sum all the step for each day with
 before we have load ggplot and plyr libraries by
 
```{r,echo=TRUE,warning=FALSE}
library(ggplot2)
library(plyr)
```


```{r}
stepday <-ddply(datos,"date", summarise,Nsteps=sum(steps))
```


and we obtain the histogram

```{r fig.width=7, fig.height=6}
qplot(x=date, y=Nsteps,data=stepday, geom="bar", stat="identity",position="dodge")

```


```{r}
mean_st <- mean(stepday$Nsteps,na.rm=TRUE)
mead_st <- median(stepday$Nsteps,na.rm=TRUE)
```
The mean of the total number of step is `r mean_st` and the median is `r mead_st`

## What is the average daily activity pattern?  
We calculated the average number of steps taken, averaged across all days with the code  

```{r}
intervalos <- ddply(datos,"interval",summarise,media_int=mean(steps,na.rm=TRUE))

```

Here we have the time series plot  of the 5-minute interval

```{r,fig.width=7, fig.height=6}
qplot(x=interval, y=media_int,data=intervalos,geom="line",stat="identity",position="dodge")

```

```{r}
Nmax <-which.max(intervalos$media_int)
Nint <- intervalos$interval[Nmax]
```
The interval that contains the maximum number of steps in average is `r Nint` 

## Imputing missing values  
```{r, echo=FALSE}
num_na <- sum(is.na(datos$steps))

```
The total number of missing values in the dataset is  `r num_na`

The strategy for filling the missing values in the dataset was take mean for the 5-minute interval and round for an integer and we use the following code  

```{r}
steps <- datos$steps
M_steps <- matrix(steps,nrow=288,ncol=61)

for (i in 1:61) {
    nind <-which(is.na(M_steps[,i]))
    if (length(nind !=0)) {
        M_steps[nind,i]=round(intervalos$media_int[nind])
    }
    
}

```

The new data created without NA's values is
```{r}
datos_no_NA <- data.frame(steps=c(M_steps),date=datos$date,interval=datos$interval)

```

and the histograms with NA's filled in

```{r, echo=FALSE}
stepday_NA <-ddply(datos_no_NA,"date", summarise,Nsteps=sum(steps))

```

```{r fig.width=7, fig.height=6}
qplot(x=date, y=Nsteps,data=stepday_NA, geom="bar", stat="identity",position="dodge")
```

```{r, echo=FALSE}
mean_NA <- mean(stepday_NA$Nsteps,na.rm=TRUE)
med_NA <- median(stepday_NA$Nsteps,na.rm=TRUE)
```

The mean of the total number of step is without NA `r mean_NA` and the median is `r med_NA`

## Are there differences in activity patterns between weekdays and weekends?

We have created a new variable in the data frame called **dayweek**  with this code  


```{r,echo=FALSE}
d_semana <- c("lunes","martes","miércoles","jueves","viernes","sábado","domingo")
day <- weekdays(datos_no_NA$date)
day <- factor(day,levels=d_semana)

datos_no_NA$dayweek <- ifelse(as.numeric(day) <6, "weekday","weekend")
w_semana <- as.factor(c("weekday","weekend"))
datos_no_NA$dayweek <- factor(datos_no_NA$dayweek,levels=w_semana)

weekday_data <- subset(datos_no_NA, dayweek=="weekday")
weekend_data <- subset(datos_no_NA, dayweek=="weekend")

intervalos_WD <- ddply(weekday_data,"interval",summarise,media_int=mean(steps,na.rm=TRUE))
intervalos_WD$dayweek="weekday"
intervalos_WE <- ddply(weekend_data,"interval",summarise,media_int=mean(steps,na.rm=TRUE))
intervalos_WE$dayweek="weekend"
intervalos_W <- rbind(intervalos_WD,intervalos_WE)

```

```{r}
library(lattice)
xyplot(media_int~interval | dayweek, data=intervalos_W,layout=c(2,1),type="l")

```
